---
title: "Vignette walkthrough of gene expression diversity and specificity across tissues and species for monogenic kidney-disease associated genes"
output: html_document
editor_options:
  markdown:
    wrap: 72
---

# R Script Following Steps from Vignette for diversity and specificity across tissues and species for monogenic kidney-disease associated genes

## Installing Dependencies

_!!! IMPORTANT NOTE !!!_

_Source the `functions/environment_setup.R` script in R environment to install `dplyr`, `BioConductor`, and `CoSIA` as packages.  Removed from the R Notebooks to reduce time and complexity of them._

```{r install-dependencies, message=FALSE, warning=FALSE}
beginning_time <- Sys.time()
library(dplyr)
library(CoSIA)
```

## Loading Monogenic Kidney Genes as sample genes to use

```{r loading-gene-data, message=FALSE, warning=FALSE}
load("../inst/extdata/proccessed/monogenic_kidney_genes.rda")
set.seed(42)
monogenic_kid_sample <- dplyr::sample_n(monogenic_kidney_genes, 20)
monogenic_kid_sample
```

## Get Tissues For Species

```{r gathering-tissue-data, message=FALSE, warning=FALSE}
output_species <- c("h_sapiens", "r_norvegicus")
map_tissues <- CoSIA::getTissues(output_species)
```

## Configuring the base CoSIA package like in vignette 1.2.4

notes:\
- Not using all the tissues in this example, just
`adult mammalian kidney & heart`\

```{r cosia-initialize}
CoSIAn_Obj_Vignette_diversity_specificity = CoSIA::CoSIAn(
  gene_set = unique(monogenic_kid_sample$Gene),
  i_species = "h_sapiens",
  o_species = c(
    "h_sapiens",
    "r_norvegicus"
  ),
  input_id = "Symbol",
  output_ids = "Ensembl_id",
  map_species = c(
    "h_sapiens",
    "r_norvegicus"
  ),
  map_tissues = c(
    "adult mammalian kidney",
    "heart"
  ),
  mapping_tool = "annotationDBI",
  ortholog_database = "HomoloGene",
  metric_type = "DS_Tissue" # across tissue
)
```

## Calculating species unique identifier mappings based on input gene symbol

```{r cosia-id-mapping}
CoSIAn_Obj_Vignette_diversity_specificity_convert <- CoSIA::getConversion(CoSIAn_Obj_Vignette_diversity_specificity)
head(CoSIA::viewCoSIAn(CoSIAn_Obj_Vignette_diversity_specificity_convert, "converted_id"))
```

## Diversity of Specificity Across Tissues

note: \# metric_type = "DF_Tissue"

```{r diversity-specificity-human-rat-tissue,  dpi=200, fig.height=13, fig.width=6 }
CoSIA_Vignette_expression_human_rat_tissue <- CoSIA::getGExMetrics(CoSIAn_Obj_Vignette_diversity_specificity_convert)
CoSIA_Vignette_diversity_specificity_tissue_plot <- CoSIA::plotCVGEx(CoSIA_Vignette_expression_human_rat_tissue)
CoSIA_Vignette_diversity_specificity_tissue_plot

end_time <- Sys.time()

time.loop <- end_time - beginning_time
cat("Time elapsed: ", round((time.loop),3), " minutes")
```
